What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-07 android-stacks.tex
new table

Throwing general exceptions is usually considered a
bad practice, because the type does not carry enough information to identify the
cause of the exceptional behaviour. In such cases, developers need to relying on
exception message which can be neither complete nor precise
~\cite{gosling2000java}.

In Table ~\ref{toptenandroid} we can observe a high prevalence of
RuntimeExceptions being thrown by the Android platform (167 out of 232).
Manually inspecting the code on which RuntimeExceptions were thrown in Android
Platform we could observe that direct instances of java.lang.RuntimeException
they were thrown for 2 main reasons: (i) when there was no reasonable way of
handling the exception, which means that regardless the kind of exception the
only possible action was to abort the application; and when wrapping exceptions
signaled by application methods. Since Android platform is heavily based on
"call backs", when the application does not handle its exceptions, they scape to
Android platform which does not have contextual information to adequately handle
the exception and simply terminates the application. What is apparently a bad
practice of signaling general exceptions in this context it was and intended
design decision.

@@

Runtime   & 1101 & 56  & 1446 & 423 & 10 & 3036  \\
Error     & 148  & 34  & 168  & 114 & 9  & 473 \\
Checked   & 115  & 241 & 139  & 176 & 12 & 683 \\
Undefined & 1    & 0   & 6    & 8   & 0  & 15  \\
Throwable & 0    & 0   & 1    & 0   & 0  & 1 \\
    \hline
All       &	1365	&	331	&	1760	&	721	&	31	&	4208	\\
    \hline
  \end{tabular}
\caption{Types of Root Exceptions.}
  \label{fig:typeroottab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Exception Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
    Runtime   & None    & 2360 & 359 \\
    Runtime   & Runtime & 560  & 182 \\
    Checked   & None    & 422  & 117 \\
    Error     & None    & 381  & 141 \\
    Runtime   & Checked & 109  & 70  \\
    Checked   & Checked & 98   & 42  \\
    Runtime   & Error   & 55   & 38  \\
    Checked   & Runtime & 22   & 12  \\
    Error     & Error   & 15   & 14  \\
    Undefined & None    & 15   & 10  \\
    Error     & Runtime & 13   & 6 \\
    Checked   & Error   & 8    & 7 \\
    Error     & Checked & 7    & 4 \\
    Throwable & Runtime & 3    & 1 \\
    Undefined & Runtime & 3    & 3 \\
    Throwable & None    & 1    & 1 \\
    Undefined & Error   & 1    & 1 \\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
\label{fig:wrappingandroid}
\end{table}

>>>>

Throwing general exceptions is  considered a
bad practice, because the type does not carry enough information to identify the
cause of the exceptional behaviour. In such cases, developers need to relying on
exception message which can be neither complete nor precise
~\cite{gosling2000java}.

In Table ~\ref{toptenandroid} we can observe a high prevalence of
RuntimeExceptions being thrown by the Android platform (167 out of 232).
Manually inspecting the code on which RuntimeExceptions were thrown in Android
Platform we could observe that direct instances of java.lang.RuntimeException
 were thrown due to two main reasons: (i) when there was no recovery action associated
to the exceptional condition, which means that regardless the kind of exception the
only possible action was to abort the application; and when translating exceptions
signaled by application methods. Since Android platform is heavily based on
"call backs", when a application throws a runtime exception (i.e., one of the life cicle methods of an Activity) and does it, the exception scapes to compotenst of the Android platform 
which do not have contextual information to adequately handle
the exception and simply terminates the application.

We could observe that although been known as a bad practice, the signaling general exceptions in this context it was an intended design decision when no specific recovery action can be defined for the exception.

@@

Runtime	&	1075	&	56	&	1446	&	374	&	10	&	2961	\\
Error	&	144	&	34	&	168	&	105	&	9	&	460	\\
Checked	&	110	&	230	&	139	&	145	&	12	&	636	\\
Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\
Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
    \hline
All		& 1330	&	320	&	1760	&	632	&	31	&	4073	\\
    \hline
  \end{tabular}
\caption{Types of Root Exceptions.}
  \label{fig:typeroottab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Root Cause} & \bfseries{Wrapper Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
     Runtime    & -   &  2360 & 359 \\

    Runtime   & Runtime & 560  & 182 \\

 Checked    &   -  & 422  & 117 \\
     Error    &   -   & 381  & 141 \\
   Checked &  Runtime   & 109  & 70  \\
    Checked   & Checked & 98   & 42  \\
   Error   &  Runtime   & 55   & 38  \\
   Runtime &  Checked   & 22   & 12  \\
    Error     & Error   & 15   & 14  \\
  Undefined    &  -  &  15   & 10  \\
   Runtime & Error     &  13   & 6 \\
  Error   & Checked   &  8    & 7 \\
    Checked &Error     &  7    & 4 \\
    Runtime &Throwable &  3    & 1 \\
   Runtime & Undefined &  3    & 3 \\
   Throwable    & -  &  1    & 1 \\
   Error   & Undefined &  1    & 1 \\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
\label{fig:wrappingandroid}
\end{table}
